---
title: "Cat Lovers"
date: "02.05.20"
output:
  html_document: default 
  pdf_document: default
---

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(DT)
```


```{r load-data, message=FALSE}
cat_lovers <- read_csv("data/cat-lovers.csv")
```

## Why isn't this working?

```{r}
cat_lovers %>%
  summarise(mean_cats = mean(number_of_cats))
```

## Why is this still not working?

```{r}
cat_lovers %>%
  summarise(mean_cats = mean(number_of_cats, na.rm = TRUE))
```

## Let's look at the data...

```{r}
glimpse(cat_lovers)
```

## Let's take another look

```{r}
cat_lovers %>%
  count(number_of_cats)
```

## Fix data entry errors

```{r}
cat_lovers %>%
  mutate(number_of_cats = case_when(
    name == "Ginger Clark" ~ 2,
    name == "Doug Bass"    ~ 3,
    TRUE                   ~ as.numeric(number_of_cats)
    )) %>%
  summarise(mean_cats = mean(number_of_cats))
```

## Now that we know what we're doing...

```{r}
cat_lovers <- cat_lovers %>%
  mutate(
    number_of_cats = case_when(
      name == "Ginger Clark" ~ "2",
      name == "Doug Bass"    ~ "3",
      TRUE                   ~ number_of_cats
      ),
    number_of_cats = as.numeric(number_of_cats)
    )
```

## Subsetting vectors


```{r}
x <- c(8, 4, 7,-1, 0, 100)
```

```{r}
x[1]
```

```{r}
x[c(1, 4)]
```

```{r}
x[c(TRUE, FALSE)]
```

## Creating data frames

```{r}
df <- tibble(x = 1:3, y = c("a", "b", "c"))
class(df)
glimpse(df)
```

## Features of data frames

```{r}
attributes(df)
class(df$x)
class(df$y)
```

## Working with tibbles in pipelines

```{r}
mean_cats <- cat_lovers %>%
  summarise(mean_cats = mean(number_of_cats))

cat_lovers %>%
  filter(number_of_cats < mean_cats) %>%
  nrow()
```

## A result of a pipeline is almost always a tibble

```{r}
mean_cats
class(mean_cats)
```

## `pull()` can be great

```{r}
mean_cats <- cat_lovers %>%
  summarise(mean_cats = mean(number_of_cats)) %>%
  pull()

cat_lovers %>%
  filter(number_of_cats < mean_cats) %>%
  nrow()
```

```{r}
mean_cats
class(mean_cats)
```

## Factors

```{r}
(x = factor(c("BS", "MS", "PhD", "MS")))
```

```{r}
glimpse(x)
```

```{r}
typeof(x)
```

## Read data in as character strings

```{r}
glimpse(cat_lovers)
```

## But coerce when plotting

```{r fig.height=6}
ggplot(cat_lovers, mapping = aes(x = handedness)) +
  geom_bar()
```

## Use `forcats` to manipulate factors

```{r fig.height=5}
cat_lovers %>%
  mutate(handedness = fct_relevel(handedness, "right", "left", "ambidextrous")) %>% #<<
  ggplot(mapping = aes(x = handedness)) +
  geom_bar()
```